\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-22 forwards-and-backwards.tex
lemma 2.8 fixup

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \smooth{\gamma} \eta(x))
    \int r(y, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,  dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(z, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{N}(x))
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}
In order to pass to a scaling limit, we will need to impose some 
conditions on the parameters of our model.

@@

{\bf Reproduction parameters:}
We assume that 
\begin{enumerate}
\item The function $F(x,m)$ satisfies 
    \begin{enumerate}
    \item $F(x,m)$ is locally Lipschitz in $m$;
    \item $F(x,m)$ is uniformly bounded above (but not necessarily below);
    \item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
    \end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
    derivatives in both arguments;
    \label{r_gamma_derivs_condition}
\item $\gamma(x,m)$ is uniformly bounded;
    \label{gamma_bounded_condition}
\item For each $f \in C_b^2(\IR^d)$,
    there is a $C_f$ such that
    $$ |\gamma(x, \smooth{\gamma} \eta(x)) \theta \int 
    (r(y, \smooth{r} \eta(y)) f(y) - r(x, \smooth{r} \eta(x)) f(x))
    q_\theta(x, dy)| \le C_f $$
    for all $x \in \IR^d$ and $\eta \in \measures$.
    Furthermore, $C_f$ only depends on the norm of $f$ and it's first two derivatives,
    i.e., 
    \[
        C_f = C(\sup_x \max_{ij} \max(f(x), \partial_{x_i} f(x), \partial_{x_i} \partial_{x_j} f(x))) .
    \] \label{gamma_B_condition}
\end{enumerate}
To keep expressions manageable, we shall also assume that 
\[
\mu_\theta(x) = r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N),
\]
that is, this expression is non-negative so that there is no need to take the maximum
	with zero in~(\ref{eqn:mu_defn}). (This is anyway true for 
	sufficiently large $\theta$.)
\end{assumptions}

We give two concrete situations in which the fourth condition is satisfied.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
purely atomic measures, with $\eta_0^N$ comprised of atoms of mass $1/N$, which is
tight 
in $\measures$ and such that there exists $\lambda>0$ for 
which $\langle e^{\lambda\|x\|}, \eta_0^N(dx)\rangle<C<\infty$
with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{proof}[of Lemma~\ref{lem:conditions on r}]
Here, we need to prove that $|\gamma(x, \eta) B^\theta_f(x, \eta)|$ is bounded,
uniformly over $x$ and $\eta$.
First suppose that
Condition~\ref{control through r} of  
Lemma~\ref{lem:conditions on r} is satisfied.
We write
\begin{align*}
    r\big(y,\rho_r*\eta(y)\big) f(y) - r\big(x,\rho_r*\eta(x)\big) f(x)
    &=
    \sum_i (y_i - x_i) \partial_{x_i} \Big( r\big(x,\rho_r*\eta(x)\big) f(x) \Big) 
    \\ &\qquad {}
    +
    \frac{1}{2} \sum_{ij} (y_i - x_i)(y_j - x_j)
        \partial_{x_i x_j} \Big( r\big(z, \smooth{r}\eta(z)\big) f(z) \Big) ,
\end{align*}
where $z = \kappa x + (1-\kappa) y$ for some $0 \le \kappa \le 1$.
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x, \rho_r*\eta(x))$
in this expression and so, provided $f$ also has uniformly bounded first and second derivatives,
we have a bound of the form
\begin{align*}
|B^\theta_f|
\leq
        K_1 \left|
            \theta \int_{\IR^d} \sum_i (y_i - x_i) q_\theta(x, dy)
        \right|
        +
        K_2 \left|
            \theta \int_{\IR^d} \sum_{ij} (y_i - x_i) (y_j - x_j)  q_\theta(x, dy)
        \right| 
\end{align*}
for suitable constants $K_1$, $K_2$.
This is bounded uniformly because
$q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$
are bounded uniformly in $x$ under Assumptions~\ref{def:model_setup}.
Since $\gamma$ is bounded as well, so is $|\gamma B^\theta_f|$.

@@

 Writing
    $K_3 = \sup_{x,m} \max_i |\partial_{x_i} f(x)r(x, m)|$
    and 
    $K_4 = \sup_{x,m} \max_{i,j} |\partial_{x_i x_j} f(x)r(x, m)|$, since
\[\big|\partial_{x_i}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_3\]
and
\[\big|\partial_{x_ix_j}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_4,\]
the first term
is bounded exactly as above. For the second, 
    \begin{multline*}
        r(x, \smooth{r}\eta(y)) - r(x, \smooth{r}\eta(x))
        \\
=
        (\smooth{r}\eta(y) - \smooth{r}\eta(x)) r'(x,\smooth{r}\eta(x))
        + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 r''(x, \overline{m}),
    \end{multline*}
    where $\overline{m}= \kappa'\smooth{r}\eta(x) +(1-\kappa') 
\smooth{r}\eta(y)$ for some $0\leq \kappa'\leq 1$, and
    we have used $r'$ and $r''$ to denote the first and second derivatives of $r(x,m)$
with respect to the second argument.
    So, writing $K_5=\|f\|_\infty\|r'\|_\infty$ and $K_6=\|f\|_\infty\|r''\|_\infty$,
    the second term in~\eqref{eqn:q_bound_split} is bounded by
    \begin{align*}
       K_5
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))
            q_\theta(x, dy)
        \right|
        +
       K_6
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
            q_\theta(x, dy)
        \right| .
    \end{align*}
Under Condition~\ref{control through gamma} of 
Assumptions~\ref{def:model_setup}, the right hand side is bounded by 
a constant times $\smooth{\gamma}\eta(x)+(\smooth{\gamma}\eta(x))^2$ and 
$m^2\gamma(m)$ is bounded, so therefore $|\gamma B^\theta_f|$ is bounded.

Finally, note that all bounds depended only on $\|f\|_\infty$,
$\max_i \|\partial_{x_i} f\|_\infty$, or 
$\max_{ij} \|\partial_{x_i} \partial_{x_j} f\|_\infty$.
\end{proof}

@@

\begin{proof}
Consider the semimartingale decomposition from equation~\eqref{eqn:eta_martingale}:
\begin{align} \label{eqn:eta_f_mgale_decomp}
    \langle f, \eta^N_t \rangle
    &=
    \langle f, \eta^N_0 \rangle
    + \int_0^t \int_{\IR^d} \big\{
        \gamma(x, \eta^N_s) B^\theta_f(x, \eta^N_s)
        + f(x) F(x, \eta^N_s)
        \big\} \eta^N_s(dx) ds
    + M^N_t(f) ,
\end{align}
where $M^N_t(f)$ is a martingale and $B^\theta_f$ is defined in \ref{generator Bf}.
First note that Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
stipulates that $|\gamma B^\theta_f|$ is uniformly bounded,
and so recalling that $F$ is bounded above, we can now conclude that
under Assumptions~\ref{def:model_setup}
$\gamma(x, \eta) B^\theta_f(x, \eta) + f(x) F(x, \eta) \le C_f $ for some $C_f$.

Now, taking expectations in~(\ref{eqn:eta_f_mgale_decomp}),
    \begin{align}
\label{bound on intfdeta}
        \IE\left[ \langle f, \eta^N_t \rangle \right]
        &\le
        \IE\left[ \langle f, \eta^N_0 \rangle \right]
        + C_f \int_0^t \IE\left[ \langle 1, \eta^N_s \rangle \right] dt .
    \end{align}
The bound~(\ref{eta_f_bound}) then follows by first applying Gronwall's
inequality in the case $f=1$, which yields
\[
	\IE\big[\langle 1,\eta_t^N\rangle\big]\leq e^{Ct}\IE\big[\langle 1,\eta_0^N\rangle\big],
\]
with $C$ independent of $N$, and then substituting 
the resulting bound on $\IE\big[\langle 1,\eta_s^N\rangle\big]$ into the
expression above.
\end{proof}

@@

\begin{proof}[Sketch]
To avoid the idea being buried in notation, we sketch the proof in the 
special case of $d=1$.
We take a sequence of test functions $f_n$ increasing to the function $f(x)=x$
and having uniformly bounded first and second derivatives.
Examining the proof of Lemma~\ref{lem:conditions on r},
so that there is a (single) $C_1$
from Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
such that $\gamma(x,\eta) B_{f_n}^\theta(x,\eta) \le C$ for all $x$, $\eta$, and $f_n$.

To avoid the idea being buried in notation, we sketch the proof in the 
special case of $d=1$ and $\DG=\Delta$.
For any $f\in C_b^\infty(\IR)$,
\begin{multline}
\label{special case first moment equation}
\IE[\langle f,\eta_t\rangle] 
    =\IE[\langle f,\eta_0\rangle]
    + \IE\Big[\int_0^t\Big\langle\gamma(x, \smooth{\gamma}\eta_s(x))
    \Delta \big(f(\cdot)r(\cdot,\smooth{r}\eta_s(\cdot))\big)(x)
    \\ \qquad 
    +F(x,\smooth{F}\eta_s(x)) f(x),
    \eta_s(dx)\Big\rangle ds\Big].
\end{multline} 
We take a sequence of test functions $f_n$ increasing to the constant function 1,
and having uniformly bounded first and second derivatives
so that there is a (single) $C_1$
from Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
such that $\gamma(x,\eta) B_{f_n}^\theta(x,\eta) \le C$ for all $x$, $\eta$, and $f_n$.
This implies that $\gamma(x,\eta) \Delta(f_n(\cdot) r(\cdot,\eta))(x) \le C_1$
for all $n$, and so in the limit, since $F$ is also bounded above,
there is a $C$ such that
\begin{align*}
    \IE[\langle 1, \eta_t\rangle]
    &=
    \IE[\langle 1, \eta_0\rangle]
    +
    \IE\Big[\int_0^t \Big\langle \gamma(x,\eta_s) \Delta r(x,\eta_s) + F(x,\eta_s) ,
    \eta_s(dx)\Big\rangle ds\Big] .
    \\ &\le
    \IE[\langle 1, \eta_0\rangle]
    +
    C \IE\Big[\int_0^t \Big\langle 1, \eta_s(dx)\Big\rangle ds\Big] .
\end{align*}
Gronwall's inequality then implies that 
$\IE[\langle 1, \eta_t(dx)\rangle]$ 
is uniformly bounded for $t\in [0,T]$, and an application of Markov's inequality
tells us that for any $\delta >0$,
$\IP[\eta_t([R,\infty])>\delta]\to 0$ as $R\to\infty$.
\end{proof}

>>>>

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \smooth{\gamma} \eta(x))
    \int r(y, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,  dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(z, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{N}(x))
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}

In order to pass to a scaling limit, we will need to impose some 
conditions on the parameters of our model.

@@

{\bf Reproduction parameters:}
We assume that 
\begin{enumerate}
\item The function $F(x,m)$ satisfies 
    \begin{enumerate}
    \item $F(x,m)$ is locally Lipschitz in $m$;
    \item $F(x,m)$ is uniformly bounded above (but not necessarily below);
    \item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
    \end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
    derivatives in both arguments;
    \label{r_gamma_derivs_condition}
\item $\gamma(x,m)$ is uniformly bounded;
    \label{gamma_bounded_condition}
\item For each $f \in C_b^2(\IR^d)$,
    there is a $C_f$ such that
    $$
        |\gamma(x, \smooth{\gamma} \eta(x)) \theta \int 
        (r(y, \smooth{r} \eta(y)) f(y) - r(x, \smooth{r} \eta(x)) f(x))
        q_\theta(x, dy)| \le C_f (1 + |f(x)|)
    $$
    for all $x \in \IR^d$ and $\eta \in \measures$.
    Furthermore, $C_f$ only depends on the norm of the first two derivatives of $f$,
    i.e., 
    \[
        C_f = C(\sup_x \sup_{\|z\|=1} \max( \sum_i z_i \partial_{x_i} f(x), \sum_{ij} z_i z_j \partial_{x_i x_j} f(x))) .
    \] \label{gamma_B_condition}
\end{enumerate}
To keep expressions manageable, we shall also assume that 
\[
\mu_\theta(x) = r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N),
\]
that is, this expression is non-negative so that there is no need to take the maximum
	with zero in~(\ref{eqn:mu_defn}). (This is anyway true for 
	sufficiently large $\theta$.)
\end{assumptions}

Since we take bounded $f$,
for most situations the bound $C_f(1+f(x))$ in condition~\ref{gamma_B_condition} above
can be safely replaced simply by $C_f$;
however, this will be useful
in certain situations where we consider a sequence of $f$ with increasing upper bounds.
We now give two concrete situations in which the fourth condition is satisfied.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
    Suppose that Assumptions~\ref{def:model_setup} hold and, 
    further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
    purely atomic measures, with $\eta_0^N$ comprised of atoms of mass $1/N$,
    which is tight in $\measures$ and such that
    $\langle |x|, \eta_0^N(dx)\rangle<C<\infty$
    with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{proof}[of Lemma~\ref{lem:conditions on r}]
Here, we need to prove that $|\gamma(x, \eta) B^\theta_f(x, \eta)|$ is bounded,
uniformly over $x$ and $\eta$.
First suppose that
Condition~\ref{control through r} of  
Lemma~\ref{lem:conditions on r} is satisfied.
\comment{BEGIN NEW VERSION}
We write
\begin{align*}
    r(y,\eta) f(y) - r(x,\eta) f(x)
&=
    r(y,\eta) (f(y) - f(x))
    + (r(y,\eta) - r(x,\eta)) f(x) 
\\ &=
    r(y,\eta) \left(
        \sum_i (y-x)_i \partial_{x_i} f(x)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} f(z_1)
    \right)
\\ &\qquad {}
    + f(x) \left(
        \sum_i (y-x)_i \partial_{x_i} r(x,\eta)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} r(z_2,\eta)
    \right)
 \\ &=
    \left( r(x,\eta) + \sum_j (y-x)_j \partial_{x_j} r(z_3,\eta) \right)
    \left(
        \sum_i (y-x)_i \partial_{x_i} f(x)
    \right)
\\ &\qquad {}
    + r(y,\eta) 
        \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} f(z_1)
\\ &\qquad {}
    + f(x) \left(
        \sum_i (y-x)_i \partial_{x_i} r(x,\eta)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} r(z_2,\eta)
    \right) ,
\end{align*}
for some $z_i = \kappa_i x + (1-\kappa_i) y$.
Integrating this against $q(x, dy)$, we get that
\begin{align*}
& \bigg| \theta \int \left(
    r(y,\eta) f(y) - r(x,\eta) f(x)
\right) q_\theta(x, dy) \bigg|
\\ &\qquad \le
    \bigg| \sum_i \left( r(x,\eta) \partial_{x_i} f(x) + f(x) \partial_{x_i} r(x,\eta) \right)
    \theta \int (y-x)_i q_\theta(x, dy) \bigg|
\\ &\qquad \qquad {} +
    \bigg| f(x) 
    \theta \int \sum_{ij} \partial_{x_i x_j} r(z_2,\eta) (y-x)_i (y-x)_j q_\theta(x, dy) \bigg|
\\ &\qquad \qquad {} +
    \bigg| \theta \int \sum_{ij} (y-x)_i (y-x)_j 
        \left(
            \partial_{x_i} f(x) \partial_{x_j} r(z_3,\eta)
            + r(y,\eta) \partial_{x_ix_j} f(z_1)
        \right) q_\theta(x, dy) \bigg|
\end{align*}
Since $q_\theta(x,dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$ are uniformly bounded,
so that $\theta \int (y-x)_i q_\theta(x,dy)$ is bounded as well.
Furthermore, a change of variables that diagonalizes $\covq(x)$
shows for any $g : \IR^d \to \IR^{d + d}$,
that if $C_g = \sup_y \sup_{\|z\| = 1} \sum_{ij} g(y)_{ij} z_i z_j$
and $\lambda_* = \sup_y \sup_{\|z\|=1} \sum_{ij} \covq(y)_{ij} z_i z_j$
then
$$ \theta \int \sum_{ij} g(y)_{ij} (y-x)_i (y-x)_j q_\theta(x, dy) \le C_g \lambda_* .  $$
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x,\eta) = r(x,\rho_r*\eta(x))$
in this expression and so, provided $f$ also has uniformly bounded first and second derivatives,
we have a bound of the form
\begin{align*}
    |B^\theta_f| \leq K_1 + K_2 |f(x)| ,
\end{align*}
for suitable constants $K_1$, $K_2$ that depend only on the derivatives of $f$.

\comment{BEGIN PREVIOUS VERSION}
We write
\begin{align*}
    r\big(y,\rho_r*\eta(y)\big) f(y) - r\big(x,\rho_r*\eta(x)\big) f(x)
    &=
    \sum_i (y_i - x_i) \partial_{x_i} \Big( r\big(x,\rho_r*\eta(x)\big) f(x) \Big) 
    \\ &\qquad {}
    +
    \frac{1}{2} \sum_{ij} (y_i - x_i)(y_j - x_j)
        \partial_{x_i x_j} \Big( r\big(z, \smooth{r}\eta(z)\big) f(z) \Big) ,
\end{align*}
where $z = \kappa x + (1-\kappa) y$ for some $0 \le \kappa \le 1$.
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x, \rho_r*\eta(x))$
in this expression and so, provided $f$ also has uniformly bounded first and second derivatives,
we have a bound of the form
\begin{align*}
|B^\theta_f|
\leq
        K_1 \left|
            \theta \int_{\IR^d} \sum_i (y_i - x_i) q_\theta(x, dy)
        \right|
        +
        K_2 \left|
            \theta \int_{\IR^d} \sum_{ij} (y_i - x_i) (y_j - x_j)  q_\theta(x, dy)
        \right| 
\end{align*}
for suitable constants $K_1$, $K_2$
that depend on $f$ only through 
$\max_i \|\partial_{x_i} f\|_\infty$ and
$\max_{ij} \|\partial_{x_i} \partial_{x_j} f\|_\infty$.
This is bounded uniformly because
$q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$
are bounded uniformly in $x$ under Assumptions~\ref{def:model_setup}.
Since $\gamma$ is bounded as well, so is $|\gamma B^\theta_f|$.
\comment{END OLD VERSION}

@@

 Writing
    $K_3 = \sup_{x,m} \max_i |\partial_{x_i} f(x)r(x, m)|$
    and 
    $K_4 = \sup_{x,m} \max_{i,j} |\partial_{x_i x_j} f(x)r(x, m)|$, since
\[\big|\partial_{x_i}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_3\]
and
\[\big|\partial_{x_ix_j}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_4,\]
the first term
is bounded exactly as above. For the second, 
    \begin{multline*}
        r(x, \smooth{r}\eta(y)) - r(x, \smooth{r}\eta(x))
        \\
=
        (\smooth{r}\eta(y) - \smooth{r}\eta(x)) r'(x,\smooth{r}\eta(x))
        + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 r''(x, \overline{m}),
    \end{multline*}
    where $\overline{m}= \kappa'\smooth{r}\eta(x) +(1-\kappa') 
\smooth{r}\eta(y)$ for some $0\leq \kappa'\leq 1$, and
    we have used $r'$ and $r''$ to denote the first and second derivatives of $r(x,m)$
with respect to the second argument.
    So, writing $K_5=\|r'\|_\infty$ and $K_6=\|r''\|_\infty$,
    the second term in~\eqref{eqn:q_bound_split} is bounded by $f(x)$ multiplied by
    \begin{align*}
       K_5
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))
            q_\theta(x, dy)
        \right|
        +
       K_6
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
            q_\theta(x, dy)
        \right| .
    \end{align*}
Under Condition~\ref{control through gamma} of 
Assumptions~\ref{def:model_setup}, this is bounded by 
a constant times $\smooth{\gamma}\eta(x)+(\smooth{\gamma}\eta(x))^2$ and 
$\sup_x m^2\gamma(x, m)$ is bounded,
so therefore $|\gamma(x,\eta) B^\theta_f(x,\eta)| \le K_7 + K_8 f(x)$,
where $K_7$ comes from $K_3$, $K_4$, and the supremum of $\gamma$,
while $K_8$ comes from $K_5$, $K_6$, and the supremum of $m^2 \gamma(x,m)$.
\end{proof}

@@

\begin{proof}
Consider the semimartingale decomposition from equation~\eqref{eqn:eta_martingale}:
\begin{align} \label{eqn:eta_f_mgale_decomp}
    \langle f, \eta^N_t \rangle
    &=
    \langle f, \eta^N_0 \rangle
    + \int_0^t \int_{\IR^d} \big\{
        \gamma(x, \eta^N_s) B^\theta_f(x, \eta^N_s)
        + f(x) F(x, \eta^N_s)
        \big\} \eta^N_s(dx) ds
    + M^N_t(f) ,
\end{align}
where $M^N_t(f)$ is a martingale and $B^\theta_f$ is defined in \ref{generator Bf}.
First note that Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
stipulates that $|\gamma B^\theta_f|$ is uniformly bounded by a contant times $1 + f$,
and so recalling that $F$ is bounded above, we conclude that
under Assumptions~\ref{def:model_setup}
$\gamma(x, \eta) B^\theta_f(x, \eta) + f(x) F(x, \eta) \le C_f (1 + f(x))$ for some $C_f$.

Now, taking expectations in~(\ref{eqn:eta_f_mgale_decomp}),
    \begin{align}
\label{bound on intfdeta}
        \IE\left[ \langle f, \eta^N_t \rangle \right]
        &\le
        \IE\left[ \langle f, \eta^N_0 \rangle \right]
        + C_f \int_0^t \IE\left[ \langle 1 + f, \eta^N_s \rangle \right] dt .
    \end{align}
The bound~(\ref{eta_f_bound}) then follows by first applying Gronwall's
inequality in the case $f=1$, which yields
\[
	\IE\big[\langle 1,\eta_t^N\rangle\big]\leq e^{Ct}\IE\big[\langle 1,\eta_0^N\rangle\big],
\]
with $C$ independent of $N$, and then substituting 
the resulting bound on $\IE\big[\langle 1,\eta_s^N\rangle\big]$ into the
expression above.
\end{proof}

@@

\begin{proof}[Sketch]
To avoid the idea being buried in notation, we sketch the proof in the 
special case of $d=1$.
We take a sequence of test functions $f_n$ increasing to the function $f(x)=x$
and having uniformly bounded first and second derivatives,
so that there is a (single) $C$
from Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
such that $\gamma(x,\eta) B_{f_n}^\theta(x,\eta) \le C(1 + f_n(x))$
for all $x$, $\eta$, and $f_n$.
Then, just as we arrived at equation~\eqref{bound on intfdeta},
\begin{align*}
    \IE[\langle f_n(x), \eta^N_t(dx) \rangle]
    &\le
    \IE[\langle f_n(x), \eta^N_0(dx) \rangle]
    +
    C \int_0^t \IE\Big[\Big\langle f_n(x), \eta^N_s(dx) \Big] \Big\rangle ds ,
\end{align*}
with the same constant for all $n$ and all $N$.
Gronwall's inequality then implies that 
$\IE[\langle f_n, \eta^N_t\rangle] \le C'$
for some $C'$ independent of $n$, $N$, and $t \in [0,T]$.
By first taking $N \to \infty$ and then $n \to \infty$,
we find that $\IE[\langle x, \eta_t(dx)\rangle] \le C'$
for $t\in [0,T]$, and an application of Markov's inequality
tells us that for any $\delta > 0$,
$\IP[\eta_t([R,\infty])>\delta]\to 0$ as $R\to\infty$.
The same argument with $f_n(x)$ increasing to $-x$ completes the proof.
\end{proof}
